---
title: "Fantasy Football EDA"
author: "Andrew Bowen"
format: html
editor: visual
---

## Fantasy EDA

Performing some basic EDA on the fantasy football data we've scraped dating back to 2000

Some questions: - Is pairing your WRs with QBs a winning strategy? - Do RBs on winning teams do better because they have more leads?

### Pairing Quarterbacks and Wide Receivers

One strategy in fantasy is to pair together a quarterback and the top wide receiver from a given team. The thinking being that if they are a dynamic passing attack you'll "double up" on any scores produced by the pair. Here we test that hypothesis

Most players only take the top WR from a team on which their QB plays, so we'll defer to taking only the top WR-QB pairing from each team in a single season

```{r message=FALSE}
library(tidyverse)
df <- read.csv("fantasy-scoring.csv")
```

For simplicity, we're going to make a few assumptions:

-   Only considering QBs and WRs (no TEs/RBs)

```{r filtering}
df <- df %>% filter(fantpos == "QB" | fantpos == "WR")

df
```

Get the top WRs and QBs for each team (in the case of injury we'll take the passing/receiving leaders for each team's season in terms of receptions/attempts, respectfully)

```{r}
all_wrs = df %>% filter(fantpos == "WR")
all_qbs = df %>% filter(fantpos == "QB")


# Get top-performers for each team each season
wrs <- all_wrs %>% 
  group_by(year, tm) %>% 
  mutate(top_receiver = max(fantasy_ppr)) %>% 
  slice(which.max(fantasy_ppr)) %>% 
  rename(wr_fantasy_ppr = fantasy_ppr) %>%
  select(player, wr_fantasy_ppr, tm, year, rank)

qbs <- all_qbs %>%
  group_by(year, tm) %>%
  mutate(top_qb = max(fantasy_ppr)) %>%
  slice(which.max(fantasy_ppr)) %>% 
  rename(qb_fantasy_ppr = fantasy_ppr) %>%
  select(player, qb_fantasy_ppr, tm, year, rank)
```

```{r}
# Join together QB and wideout data
dat <- qbs %>% left_join(wrs, by=c("tm", "year"))
```

First let's just see the comparison between a QB's fantasy points (PPR scoring) and their top Wide Receivers. We'd expect a positive relationship here

```{r}
# Now let's plot the WR production as a function of QB production - this should be linear
dat %>% ggplot(aes(x=qb_fantasy_ppr, y = wr_fantasy_ppr)) + geom_point()
```

Nothing too surprising here, there seems to be a slight positive correlation between the fantasy scoring for a QB and their WR

```{r}
cor(dat$qb_fantasy_ppr, dat$wr_fantasy_ppr)

```

What's more interesting is if a team's overall record plays in here. There are two schools of thought, one that
